---
title: "CCBER Clethodim Experiment"
author: "Joanna Tang"
date: "9/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}

# Load packages, read in data

library(tidyverse)
library(janitor)
library(plotrix)



clethodim_seeds <- read_csv("clethodim_seeds.csv") %>% filter(treatment != "NA")
clethodim_vigor <- read_csv("clethodim_vigor.csv") %>% mutate(date = as.Date(date, format = "%m/%d/%y"))


```

# Do clethodim herbicide treatments affect the fitness of Nipomo lupine?

### H1: Because clethodim is a grass-specific herbicide, application (regardless of treatment -- direct or aerial -- or seedlings stage) should have no significant effect on the viable seed production of Nipomo lupine.
#### Result: No significant effect of any treatment, regardless of seedling stage.  H1 supported.

```{r message = FALSE, echo = FALSE, warning = FALSE}

# H1: Because clethodim is a grass-specific herbicide, application (regardless of treatment -- direct or aerial -- or seedlings stage) should have no significant effect on the viable seed production of Nipomo lupine.

## Dataframe of viable seeds/pod, % viable seeds
seed_ratios <- clethodim_seeds %>% 
  mutate(viable_seeds = as.numeric(viable_seeds)) %>% 
  mutate(viable_per_pod = viable_seeds/seed_pods) %>% 
  mutate(percent_viable = (viable_seeds)/(viable_seeds+non_viable_seeds))

## Exploratory graphs -- checking for normality
pods_hist <- ggplot(seed_ratios, aes(x = seed_pods)) +
  geom_histogram()
#pretty normal
pods_qq <- ggplot(seed_ratios, aes(sample = seed_pods)) +
  geom_qq()
#pretty normal
total_seeds_hist <- ggplot(seed_ratios, aes(x = viable_seeds)) +
  geom_histogram()
#mostly normal...
total_seeds_qq <- ggplot(seed_ratios, aes(sample = viable_seeds)) +
  geom_qq()
#mosly normal...
seeds_per_pod_hist <- ggplot(seed_ratios, aes(x = viable_per_pod)) +
  geom_histogram()
#pretty normal
pods_qq <- ggplot(seed_ratios, aes(sample = viable_per_pod)) +
  geom_qq()
#pretty normal
percent_viable_hist <- ggplot(seed_ratios, aes(x = percent_viable)) +
  geom_histogram()
#skewed left
percent_viable_qq <- ggplot(seed_ratios, aes(sample = percent_viable)) +
  geom_qq()
#skewed left


## Dataframe of means and standard errors
summary_seed_ratios <- seed_ratios %>% 
  group_by(treatment, treatment_number) %>% 
  drop_na() %>% 
  summarize(mean_seeds = mean(viable_seeds),
            se_seeds = std.error(viable_seeds),
            mean_per_pod = mean(viable_per_pod),
            se_per_pod = std.error(viable_per_pod),
            mean_pods = mean(seed_pods),
            se_pods = std.error(seed_pods))

## Dataframe of means and standard errors, by seedling stage
summary_seed_ratios_stage <- seed_ratios %>% 
  group_by(treatment, treatment_number, seedling_stage) %>% 
  drop_na() %>% 
  summarize(mean_seeds = mean(viable_seeds),
            se_seeds = std.error(viable_seeds),
            mean_per_pod = mean(viable_per_pod),
            se_per_pod = std.error(viable_per_pod),
            mean_pods = mean(seed_pods),
            se_pods = std.error(seed_pods))

### Column graph of viable seeds per pod, by treatment, by seedling stage
seeds_per_pod_stage_col <- summary_seed_ratios_stage %>% 
  ggplot(aes(x = treatment, y = mean_per_pod)) +
  geom_col(aes(fill = seedling_stage)) +
  geom_errorbar(aes(x = treatment, ymin = mean_per_pod-se_per_pod, ymax = mean_per_pod+se_per_pod, width = .5)) +
  facet_wrap(~seedling_stage) +
  scale_fill_manual(values = c("darkgreen", "chartreuse4"), name = "Seedling Stage") +
  labs(title = "Viable seeds per pod, by seedling stage", x = "Treatment", y = "Average number of seeds", caption = "No sig difference (ANOVA p > .05)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
scale_x_discrete(labels = c("Aerial Spray", "Control", "Direct Spray"))
seeds_per_pod_stage_col

### ANOVAs of viable seeds per pod by treatment, by seedling stage
flowering_seed_ratios <- seed_ratios %>% 
  filter(seedling_stage == "flowering")
flowering_seeds_per_pod_anova <- aov(viable_per_pod ~ treatment, data = flowering_seed_ratios)
#summary(flowering_seeds_per_pod_anova)
#p = .928
germinating_seed_ratios <- seed_ratios %>% 
  filter(seedling_stage != "flowering")
germinating_seeds_per_pod_anova <- aov(viable_per_pod ~ treatment, data = germinating_seed_ratios)
#summary(germinating_seeds_per_pod_anova)
#p = .41

### Column graph of viable seeds per pod, by treatment
seeds_per_pod_col <- summary_seed_ratios %>% 
  ggplot(aes(x = treatment, y = mean_per_pod)) +
  geom_col() +
  geom_errorbar(aes(x = treatment, ymin = mean_per_pod-se_per_pod, ymax = mean_per_pod+se_per_pod, width = .5)) +
  labs(title = "Viable seeds per pod", x = "Treatment", y = "Average number of seeds", caption = "No sig difference (ANOVA p = 0.42)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
scale_x_discrete(labels = c("Aerial Spray", "Control", "Direct Spray"))
seeds_per_pod_col

### ANOVA of viable seeds per pod, by treatment
seeds_per_pod_anova <- aov(viable_per_pod ~ treatment, data = seed_ratios)
#summary(seeds_per_pod_anova)
#p = .42

### Column graph of pods per plant, by treatment, by seedling stage
pods_per_plant_stage_col <- summary_seed_ratios_stage %>% 
  ggplot(aes(x = treatment, y = mean_pods)) +
  geom_col(aes(fill = seedling_stage)) +
  geom_errorbar(aes(x = treatment, ymin = mean_pods-se_pods, ymax = mean_pods+se_pods, width = .5)) +
  facet_wrap(~seedling_stage) +
  scale_fill_manual(values = c("darkgreen", "chartreuse4"), name = "Seedling Stage") +
  labs(title = "Pods per plant, by seedling stage", x = "Treatment", y = "Average number of pods per plant", caption = "No sig difference (ANOVA p > 0.1)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
scale_x_discrete(labels = c("Aerial Spray", "Control", "Direct Spray"))
pods_per_plant_stage_col

### ANOVAs of pods per plant by treatment, by seedling stage
flowering_seeds_per_pod_anova <- aov(seed_pods ~ treatment, data = flowering_seed_ratios)
#summary(flowering_seeds_per_pod_anova)
#p = .129
germinating_seeds_per_pod_anova <- aov(viable_per_pod ~ treatment, data = germinating_seed_ratios)
#summary(germinating_seeds_per_pod_anova)
#p = .41

### Column graph of pods per plant, by treatment
pods_per_plant_col <- summary_seed_ratios %>% 
  ggplot(aes(x = treatment, y = mean_pods)) +
  geom_col() +
  geom_errorbar(aes(x = treatment, ymin = mean_pods-se_pods, ymax = mean_pods+se_pods, width = .5)) +
  labs(title = "Pods per plant", x = "Treatment", y = "Average number of pods per plant", caption = "No sig difference (ANOVA p = 0.145)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
scale_x_discrete(labels = c("Aerial Spray", "Control", "Direct Spray"))
pods_per_plant_col

### ANOVA of  pods per plant, by treatment
pods_per_plant_anova <- aov(seed_pods ~ treatment, data = seed_ratios)
#summary(pods_per_plant_anova)
#p = .145

### Column graph of seeds per plant, by treatment, by seedling stage
seeds_per_plant_stage_col <- summary_seed_ratios_stage %>% 
  ggplot(aes(x = treatment, y = mean_seeds)) +
  geom_col(aes(fill = seedling_stage)) +
  geom_errorbar(aes(x = treatment, ymin = mean_seeds-se_seeds, ymax = mean_seeds+se_seeds, width = .5)) +
  facet_wrap(~seedling_stage) +
  scale_fill_manual(values = c("darkgreen", "chartreuse4"), name = "Seedling Stage") +
  labs(title = "Viable seeds per plant, by seedling stage", x = "Treatment", y = "Average number of seeds per plant", caption = "No sig difference (ANOVA p > 0.1)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
scale_x_discrete(labels = c("Aerial Spray", "Control", "Direct Spray"))
seeds_per_plant_stage_col

### ANOVAs of seeds per plant by treatment, by seedling stage
flowering_seeds_anova <- aov(viable_seeds ~ treatment, data = flowering_seed_ratios)
#summary(flowering_seeds_anova)
#p = .126
germinating_seeds_anova <- aov(viable_seeds ~ treatment, data = germinating_seed_ratios)
#summary(germinating_seeds_anova)
#p = .313

### Column graph of seeds per plant, by treatment
seeds_per_plant_col <- summary_seed_ratios %>% 
  ggplot(aes(x = treatment, y = mean_seeds)) +
  geom_col() +
  geom_errorbar(aes(x = treatment, ymin = mean_seeds-se_seeds, ymax = mean_seeds+se_seeds, width = .5)) +
  scale_fill_manual(values = c("darkgreen", "chartreuse4"), name = "Seedling Stage") +
  labs(title = "Viable seeds per plant", x = "Treatment", y = "Average number of seeds per plant", caption = "No sig difference (ANOVA p = 0.0843)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
scale_x_discrete(labels = c("Aerial Spray", "Control", "Direct Spray"))
seeds_per_plant_col

### ANOVA of  seeds per plant, by treatment
seeds_per_plant_anova <- aov(viable_seeds ~ treatment, data = seed_ratios)
#summary(seeds_per_plant_anova)
#p = .0843

```




### H2: Because clethodim is a grass-specific herbicide, application (regardless of treatment -- direct or aerial -- or seedling stage) should have no significant effect on the vigor of Nipomo lupine.
#### Result: No significant effect of any treatment, regardless of seedling stage.  H2 supported.

```{r message = FALSE, warning = FALSE, echo = FALSE}

# H2: Because clethodim is a grass-specific herbicide, application (regardless of treatment -- direct or aerial -- or seedling stage) should have no significant effect on the vigor of Nipomo lupine.

# Dataframe calculating average vigor
summary_vigor <- clethodim_vigor %>% 
  group_by(date, treatment, seedling_stage) %>% 
  summarize(mean_vigor = mean(vigor),
            se_vigor = std.error(vigor))
  
# Dataframe calculating % at each vigor
vigor_proportions <- clethodim_vigor %>% 
  count(date, seedling_stage, vigor, treatment) %>% 
  group_by(date, treatment, seedling_stage, vigor) %>% 
  summarize(proportion = sum(n)/30)


# Scatterplot of vigor over time
vigor_scatter <- clethodim_vigor %>% 
  ggplot(aes(x = date, y = vigor)) +
  geom_point(aes(color = seedling_stage)) +
  geom_smooth(method = lm, aes(x = date, group = seedling_stage), se = TRUE, size = .5) +
  facet_wrap(~treatment) +
  labs(title = "Plant vigor over time", x = "Date", y = "Vigor Index", caption = "Flowering seedlings experienced decrased vigor over time") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(expand = c(0,0), limits = c(0.5, 4.5))
vigor_scatter

# Stacked column graph of vigor proportion over time
vigor_col <- vigor_proportions %>% 
  ggplot(aes(x = date, y = proportion, fill = as.factor(vigor))) +
  geom_col(position = "fill") +
  facet_grid(seedling_stage ~ treatment) +
  scale_fill_manual(values = c("chartreuse4", "yellowgreen", "yellow2", "coral", "brown3", "tan3", "tan4"), name = "Vigor") +
  labs(title = "Vigor proportions over time", x = "Date", y = "Proportion", caption = "Sig more mortality for flowering seedlings regardless of treatment (chi-squared p << 0.01), \n no sig difference between treatments") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
vigor_col


## Chi-squared test of vigor proportions at 6/7/20 & 6/25/20, between seedling stages
vigor_seedling_chi_squared_matrix <- clethodim_vigor %>% 
  filter(final_observation == "final") %>% 
  count(seedling_stage, vigor) %>% 
  spread(vigor, n) %>% # tell the fxn which column to spread by and what to fill the obs with
  select(-seedling_stage)
vigor_seedling_chi_squared_matrix[is.na(vigor_seedling_chi_squared_matrix)] <- 0
rownames(vigor_seedling_chi_squared_matrix) <- c("Flowering", "Newly Germinated")
vigor_seedling_chi_squared <- chisq.test(vigor_seedling_chi_squared_matrix)
#vigor_seedling_chi_squared
#p << .001 (regardless of including 3 herbivory mortalities)

## Chi-squared test of vigor proportions at 6/7/20, for newly germinated seedlings between treatments
vigor_germinated_treatment_chi_squared_matrix <- clethodim_vigor %>% 
  filter(final_observation == "final") %>% 
  filter(seedling_stage != "flowering") %>% 
  count(treatment, vigor) %>% 
  spread(vigor, n) %>% # tell the fxn which column to spread by and what to fill the obs with
  select(-treatment)
vigor_germinated_treatment_chi_squared_matrix[is.na(vigor_germinated_treatment_chi_squared_matrix)] <- 0
rownames(vigor_germinated_treatment_chi_squared_matrix) <- c("Aerial", "Control", "Direct")
vigor_germinated_treatment_chi_squared <- chisq.test(vigor_germinated_treatment_chi_squared_matrix)
#vigor_germinated_treatment_chi_squared
#p = .1461

## Chi-squared test of vigor proportions at 6/25/20, for flowering seedlings between treatments
vigor_flowering_treatment_chi_squared_matrix <- clethodim_vigor %>% 
  filter(final_observation == "final") %>% 
  filter(seedling_stage == "flowering") %>% 
  count(treatment, vigor) %>% 
  spread(vigor, n) %>% # tell the fxn which column to spread by and what to fill the obs with
  select(-treatment)
vigor_flowering_treatment_chi_squared_matrix[is.na(vigor_flowering_treatment_chi_squared_matrix)] <- 0
rownames(vigor_flowering_treatment_chi_squared_matrix) <- c("Aerial", "Control", "Direct")
vigor_flowering_treatment_chi_squared <- chisq.test(vigor_flowering_treatment_chi_squared_matrix)
#vigor_flowering_treatment_chi_squared
#p = .03567

```
